---
title: 'Assignment 3: Santa Barbara Urchins'
author: "Katie Pietrucha"
date: "10/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(janitor)
```
# Overview

In this task, you will use version control, project management, and data wrangling, analysis & visualization skills to explore red and purple urchins recorded by Santa Barbara Coastal Long Term Ecological Research program divers, while practicing version control and project management.

**Data:** Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

# Read in the data and convert column names to lower snake case using `clean_names`.

```{r}
urchins_data <- read_csv(here("data", "urchins.csv")) %>%
  clean_names
```

# Section 1: Size comparisons, purple and red urchins

```{r}
ggplot(data = urchins_data, aes(x = size, fill = common_name)) +
  scale_fill_manual(values = c("purple", "red")) +
  geom_histogram(bins = 10) +
  facet_wrap(~ common_name) +
  theme_bw() +
  labs(x = "Size (cm)", y = "Number of urchins") +
  ggtitle("Distribution of urchin sizes by species") +
  theme(plot.title = element_text(hjust = 0.5))
```

